Forecasting the United States 2024 Presidential Election

### 2024-11-04 paper/paper.qmd
fixed up

To predict state level support and calculate the expected Electoral Votes (EV) for each candidate in the upcoming US election, we employed a Bayesian linear regression model. The model is mathematically specified as follows:

@@

All analyses were conducted using R (@citeR) with the following key packages:

@@

## Deep Red/Blue State Voting History {#sec-voting-hist}

```{r}
#| fig-cap: Voting History of the 10 States That do not Have Poll Data
#| echo: false
#| warning: false
#| message: false

@@

# -----------------------------
# Step 3: Reshape the Data for a Compact Table
# -----------------------------

# Pair the states into two columns per row
# We'll split the data into two halves
df1 <- df[1:5, ]  # First 5 states
df2 <- df[6:10, ] # Next 5 states

@@

# -----------------------------
# Step 4: Apply Cell-wise Coloring Based on Party
# -----------------------------

# Function to assign colors based on party
assign_color <- function(party) {
  ifelse(party == "Rep", "#DD1717", "#0F4392")
}

# Apply cell_spec to each cell that needs coloring
compact_df <- compact_df %>%
  mutate(
    State_1 = cell_spec(State_1, background = assign_color(Party_1), color = "white"),
    Year_1 = cell_spec(Year_1, background = assign_color(Party_1), color = "white"),
    State_2 = cell_spec(State_2, background = assign_color(Party_2), color = "white"),
    Year_2 = cell_spec(Year_2, background = assign_color(Party_2), color = "white")
  )

# -----------------------------
# Step 5: Create the Styled Compact Table with Note
# -----------------------------

# Create and style the table
styled_table <- compact_df %>%
  select(State_1, Year_1, State_2, Year_2) %>%
  kable("html", escape = FALSE, align = "c",
        col.names = c("State", "Last Time Voted for a Different Party",
                      "State", "Last Time Voted for a Different Party")) %>%
  kable_styling("striped", full_width = FALSE, position = "left") %>%
  # Header styling
  row_spec(0, bold = TRUE, background = "#D3D3D3") %>%
  # Add a footnote below the table
  add_footnote("Note: Data collected only went back to 1976, some states may have been voting for the same party farther back than 1976.",
  )

# Display the table
styled_table
```
## State By State Win {#sec-state-wins}

>>>>

All model construction and analysis was conducted using R (@citeR) with the following key packages:

To predict state level support and calculate the expected Electoral Votes (EV) for each candidate in the upcoming US election, we employed a Bayesian linear regression model. The model is specified as follows:

@@

@@

## Deep Red/Blue State Voting History {#sec-voting-hist}
```{r}
#| fig-cap: Number of Wins in 1000 Simulations by State
#| echo: false
#| warning: false
#| message: false

#| fig-cap: Voting History of the 10 States That do not Have Poll Data
#| echo: false
#| warning: false
#| message: false

library(kableExtra)
library(dplyr)

@@

# Pair the states into two columns per row
df1 <- df[1:5, ]  # First 5 states
df2 <- df[6:10, ] # Next 5 states

@@

# Function to assign color names based on party
assign_color <- function(party) {
  ifelse(party == "Rep", "#DD1717", "#0F4392")
}

# Apply cell_spec to each cell that needs coloring
compact_df <- compact_df %>%
  mutate(
    State_1 = cell_spec(State_1, background = assign_color(Party_1), color = "white", format = "latex", escape = FALSE),
    Year_1 = cell_spec(Year_1, background = assign_color(Party_1), color = "white", format = "latex", escape = FALSE),
    State_2 = cell_spec(State_2, background = assign_color(Party_2), color = "white", format = "latex", escape = FALSE),
    Year_2 = cell_spec(Year_2, background = assign_color(Party_2), color = "white", format = "latex", escape = FALSE)
  )

# Create and style the table
styled_table <- compact_df %>%
  select(State_1, Year_1, State_2, Year_2) %>%
  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "c",
        col.names = c("State", "Last Time Voted for a Different Party",
                      "State", "Last Time Voted for a Different Party")) %>%
  kable_styling(latex_options = c("striped"), full_width = FALSE, position = "left") %>%
  row_spec(0, bold = TRUE) %>%
  footnote(general = " Data collected only went back to 1976, some states may have been voting for the same party farther back than 1976.")

# Display the table
styled_table

```
## State By State Win {#sec-state-wins}
